Trans-Dimensional Bayesian Inference for Gravitational Lens Substructures

### 2015-09-11 Papers/ms.tex
SPH
More honest abstract

\begin{abstract}
We introduce a Bayesian solution to the problem of inferring the density
profile of strong gravitational lenses when the lens galaxy may contain
multiple dark or faint substructures. The source and lens models are based on
a superposition of an unknown number of non-negative basis functions
(or ``blobs'') whose form was chosen with speed as a primary criterion.
The prior distribution for the blobs' properties is specified hierarchically,
so the mass function of substructures is a natural output of the method.
We use reversible
jump Markov Chain Monte Carlo (MCMC) within Diffusive Nested Sampling (DNS) to
sample the posterior distribution and evaluate the marginal likelihood of the
model, including the summation over the unknown
number of blobs in the source and the lens.
We demonstrate the method on {\revisions two
simulated data sets: one with a single
substructure, and one with ten}. We also
apply the method to the g-band
image of the ``Cosmic Horseshoe'' system, and find some
hints of potential substructures. However, we caution that such results could
also be caused by misspecifications in the model (such as the shape of
the smooth lens component or the point spread function),
which are difficult to guard against in full generality.
\end{abstract}

@@

\subsection{The Source}
The surface brightness profile of the source is assumed to be composed of
a sum of a finite number of ``blobs'', or basis functions, in order to
allow some flexibility while incorporating prior information about the
non-negativity of surface brightness, and the spatial correlation expected
in real surface brightness profiles.
For computational speed, we choose the following
functional form for the surface brightness profile
of a single blob centered at the origin:
\begin{eqnarray}
f(x, y) &=& \left\{
\begin{array}{lr}
\frac{2A}{\pi w^2}\left(1 - \frac{r^2}{w^2}\right), & r \leq w\\
0, & \textnormal{otherwise}
\end{array}
\right.
\end{eqnarray}
where $r = \sqrt{x^2 + y^2}$, $w$ is the width of the blob, and $A$ is the
total flux of the blob (i.e. the integral of the surface brightness over the
entire domain).
These basis functions are inverted paraboloids, which are faster to evaluate
than gaussians, since they do not
contain an exponential function. In addition, the finite support means that
each blob will evaluate to zero over a large fraction of the domain which
confers an additional speed advantage.

@@

>>>>

\begin{abstract}
We introduce a Bayesian solution to the problem of inferring the density
profile of strong gravitational lenses when the lens galaxy may contain
multiple dark or faint substructures. The source and lens models are based on
a superposition of an unknown number of non-negative basis functions
(or ``blobs'') whose form was chosen with speed as a primary criterion.
The prior distribution for the blobs' properties is specified hierarchically,
so the mass function of substructures is a natural output of the method.
We use reversible
jump Markov Chain Monte Carlo (MCMC) within Diffusive Nested Sampling (DNS) to
sample the posterior distribution and evaluate the marginal likelihood of the
model, including the summation over the unknown
number of blobs in the source and the lens.
We demonstrate the method on {\revisions two
simulated data sets: one with a single
substructure, and one with ten}. We also
apply the method to the g-band
image of the ``Cosmic Horseshoe'' system, and find {\revisions evidence for more
than zero substructures}. However, {\revisions these have large spatial
extent and probably only point to misspecifications in the model}
(such as the shape of
the smooth lens component or the point spread function),
which are difficult to guard against in full generality.
\end{abstract}

@@

\subsection{The Source}
The surface brightness profile of the source is assumed to be composed of
a sum of a finite number of ``blobs'', or basis functions, in order to
allow some flexibility while incorporating prior information about the
non-negativity of surface brightness, and the spatial correlation expected
in real surface brightness profiles.
For computational speed, we choose the following
functional form for the surface brightness profile
of a single blob centered at the origin:
\begin{eqnarray}
f(x, y) &=& \left\{
\begin{array}{lr}
\frac{2A}{\pi w^2}\left(1 - \frac{r^2}{w^2}\right), & r \leq w\\
0, & \textnormal{otherwise}
\end{array}
\right.
\end{eqnarray}
where $r = \sqrt{x^2 + y^2}$, $w$ is the width of the blob, and $A$ is the
total flux of the blob (i.e. the integral of the surface brightness over the
entire domain).
These basis functions are inverted paraboloids, which are faster to evaluate
than gaussians, since they do not
contain an exponential function. In addition, the finite support means that
each blob will evaluate to zero over a large fraction of the domain which
confers an additional speed advantage. {\revisions More conventional choices such as
gaussians are possible, as are alternative compact basis functions, perhaps
inspired by the smoothed particle hydrodynamics literature
\citep{2012MNRAS.425.1068D}.}

@@

{\revisions
\bibitem[\protect\citeauthoryear{Dehnen 
\& Aly}{2012}]{2012MNRAS.425.1068D} Dehnen W., Aly H., 2012, MNRAS, 425, 1068
}
